---
title: "Teacher Surveys"
author: 
output: 
  flexdashboard::flex_dashboard:
    theme: yeti
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(rio)
library(here)
library(tidyverse)
library(DT)

#Colors
#1= "#268bd2" 
#2= "#dc322f"
#3= "#859900"
#4= "#073642"
```

```{r wave1_data}

wave_1 <- import(here::here("_posts", 
                                   "2019-05-16-consequential-validity-teacher-surveys", 
                                   "data", 
                                   "wave_1.csv")) %>% 
  as_tibble()

header_1 <- import(here::here("_posts", 
                                   "2019-05-16-consequential-validity-teacher-surveys", 
                                   "data", 
                                   "header_1.csv")) %>% 
  as_tibble()

```

```{r wave2_data}

wave_2 <- import(here::here("_posts", 
                                   "2019-05-16-consequential-validity-teacher-surveys", 
                                   "data", 
                                   "wave_2.csv")) %>% 
  as_tibble()

header_2 <- import(here::here("_posts", 
                                   "2019-05-16-consequential-validity-teacher-surveys", 
                                   "data", 
                                   "header_2.csv")) %>% 
  as_tibble()

```

```{r wave3_data}

wave_3 <- import(here::here("_posts", 
                                   "2019-05-16-consequential-validity-teacher-surveys", 
                                   "data", 
                                   "wave_3.csv")) %>% 
  as_tibble()

header_3 <- import(here::here("_posts", 
                                   "2019-05-16-consequential-validity-teacher-surveys", 
                                   "data", 
                                   "header_3.csv")) %>% 
  as_tibble()

```

Sidebar {.sidebar}
=====================================
The data and comments presented here are from the teacher particapants of our [Consequential Validity Study](https://jnese.github.io/core-blog/posts/2019-05-14-consequential-validity-study-procedures/), a longitudinal study in each of the 2017-18 and 2018-19 school years, with four waves of assessments during each year. 

After each assessment wave, participating teachers responded to brief surveys with questions about their opinions about traditional oral reading fluency (ORF) assessments, the CORE system, the assessment environment, or the study itself. Survey items were different across the four waves, and the 2018-19 surveys were mostly repeats of the 2017-18 survey.


Traditional ORF Assessment {data-icon="fa-edit"}
===================================== 

Row {data-height=300} 
-----------------------------------------------------------------------

### **`r header_2$q7`** 

```{r wave2_q7}
wave_2 %>% 
  ggplot(aes(q7)) +
  geom_bar(fill = "#268bd2") + 
  theme_bw() +
  labs(x = "",
       y = "") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 50, by = 10)) +
  theme(text = element_text(size = 30))
```

### **`r header_2$q13`** 

```{r wave2_q13}
wave_2 %>% 
  filter(q13 != "") %>% 
  ggplot(aes(q13)) +
  geom_bar(fill = "#dc322f") + 
  theme_bw() +
  labs(x = "",
       y = "") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 50, by = 10)) +
  theme(text = element_text(size = 30))
```

Row {data-height=300} 
-----------------------------------------------------------------------

### **`r header_2$q10`** {data-width=300}

```{r wave2_q10}
wave_2 %>% 
  ggplot(aes(q10)) +
  geom_bar(fill = "#859900") + 
  theme_bw() +
  labs(x = "",
       y = "") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 50, by = 10)) +
  theme(text = element_text(size = 30))
```

### wave2_q10 comments {.no-title}

```{r wave2_q10comments}
q10_table2 <- wave_2 %>% 
  select(year, q10, q10_comments) %>%
  filter(q10_comments != "") %>% 
  arrange(factor(q10, levels = c("Yes", "No")), desc(q10)) %>% 
    rename("Year" = year,
         "Rating" = q10,
         "Comments" = q10_comments)

datatable(q10_table2,
          rownames = FALSE,
          caption = header_2$q10,
          options = list(pageLength = 10, dom = 'tip'))
          #options = list(dom = 't'))
```


Row {data-height=300}
-----------------------------------------------------------------------

### **`r header_2$q2`** {data-width=300}

```{r wave2_q2}
wave_2 %>% 
  ggplot(aes(q2)) +
  geom_bar(fill = "#073642") + 
  theme_bw() +
  labs(x = "",
       y = "") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 100, by = 20)) +
  theme(text = element_text(size = 30))

```

### wave2_q2 comments {.no-title} 

```{r wave2_q2comments}
q2_table <- wave_2 %>% 
  select(year, q2, q2_comments) %>%
  filter(q2_comments != "") %>% 
  arrange(factor(q2, levels = c("Small groups", "Whole class")), desc(q2)) %>% 
    rename("Year" = year,
         "Rating" = q2,
         "Comments" = q2_comments)

datatable(q2_table,
          rownames = FALSE,
          caption = header_2$q2,
          options = list(pageLength = 10, dom = 'tip'))
#          options = list(dom = 't'))
```

CORE Testing Procedures {data-icon="fa-headset"}
===================================== 

### **`r header_1$q2`** 

```{r wave1_q2}
wave_1 %>% 
  ggplot(aes(q2)) +
  geom_bar(fill = "#268bd2") + 
  theme_bw() +
  labs(x = "",
       y = "") +
  scale_y_continuous(breaks = seq(0, 80, by = 10)) +
  theme(text = element_text(size = 20)) + 
  coord_flip()

```

### **`r header_1$q3`** 

```{r wave1_q3}
wave_1 %>% 
  mutate(q3 = factor(q3,
                     levels = c("Very uncomfortable", "Somewhat uncomfortable", "Neutral", "Somewhat comfortable", "Very comfortable"))) %>% 
  ggplot(aes(q3)) +
  geom_bar(fill = "#dc322f") + 
  theme_bw() +
  labs(x = "",
       y = "") +
  scale_x_discrete(breaks = unique(wave_1$q3), 
    labels = c("Somewhat \ncomfortable", "Very \ncomfortable", "Very \nuncomfortable", "Somewhat \nuncomfortable", "Neutral")) +
   scale_y_continuous(breaks = seq(0, 60, by = 10)) +
  theme(text = element_text(size = 20)) + 
  coord_flip()
```

Comments {data-icon="fa-comment"}
===================================== 


Row {data-height=350}
-----------------------------------------------------------------------

### **`r header_1$q7`**

```{r wave1_q7, fig.width = 10, fig.asp = .66}
wave_1 %>% 
  ggplot(aes(q7)) +
  geom_bar(fill = "#268bd2") + 
  theme_bw() +
  labs(x = "",
       y = "") +
  coord_flip() +
  scale_y_continuous(breaks = seq(0, 60, by = 10)) +
  theme(text = element_text(size = 20))

ggsave(plot = last_plot(), filename = here::here("_posts", "2019-05-16-consequential-validity-teacher-surveys", "CORE_v_ORF.png"))

```

Row {.tabset}
-----------------------------------------------------------------------

### CORE vs. Traditional ORF

```{r wave1_q7comments}
q7_table <- wave_1 %>% 
  select(year, q7, q7_comments) %>%
  filter(q7_comments != "") %>% 
  arrange(factor(q7, levels = c("CORE", "Traditional ORF")), desc(q7)) %>% 
    rename("Year" = year,
         "Rating" = q7,
         "Comments" = q7_comments)

datatable(q7_table,
          rownames = FALSE,
          caption = pull(header_1, q7))#,
#          options = list(dom = 't'))
```

### ORF Imrpovements/Barriers

```{r wave3_q10}
q10_table3 <- wave_3 %>% 
  select(year, q10) %>%
  filter(q10 != "") %>%  
    rename("Year" = year,
           "Comments" = q10)

datatable(q10_table3,
          rownames = FALSE,
          caption = pull(header_3, q10))#,
#          options = list(dom = 't'))
```

### General Comments

```{r wave1-2_q9}
q9_table1 <- wave_1 %>% 
  filter(is.na(q9_keep),
         q9 != "") %>% 
  select(year, q9) %>%
  mutate(Wave = 1) %>% 
  select(year, Wave, q9) %>% 
    rename("Year" = year,
           "Comments/Feedback" = q9)

q9_table2 <- wave_2 %>%
  filter(is.na(q9_keep),
         q9 != "") %>%
  select(year, q9) %>%
  mutate(Wave = 2) %>% 
  select(year, Wave, q9) %>% 
    rename("Year" = year,
           "Comments/Feedback" = q9)

q9_table <- bind_rows(q9_table1, q9_table2)

datatable(q9_table,
          rownames = FALSE,
          caption = "General comments and feedback.")#,
 #         options = list(dom = 't'))
```